---
title: "five38_mlb_HTML"
output: html_document
date: "2025-02-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Week 1 - Assignment: MLB ELO Dataset from 538.com

For this first assignment, I am creating a dataframe from a 538 data repository and posting it to my RPUBS account.


### Analysis objective

My objective is to isolate the playoff games from this dataset.


### Loading & displaying dataframe

Load the data into a dataframe and display the internal structure of this dataframe object.

```{r}
library(tidyverse)
five38_mlb <- read_csv("https://projects.fivethirtyeight.com/mlb-api/mlb_elo.csv")
str(five38_mlb)
```


### Creating a subset focusing on the Playoff Rounds.

```{r}
five38_mlb_subset <- five38_mlb[c("date", "season", "playoff", "team1", "team2", "score1", "score2")]
```


### Reducing the observations of the subset to only actual playoff games.

```{r}
mlb_playoff_games <- na.omit(five38_mlb_subset)
str(mlb_playoff_games)
```


### Renaming column names to improve dataframe readability.

```{r}
mlb_playoff_games <- mlb_playoff_games |> 
    rename(
        Date_of_game = 'date',
        Year_of_season = 'season',
        Playoff_Round = 'playoff',
        Home_Team = 'team1',
        Away_Team = 'team2',
        Home_Team_Score = 'score1',
        Away_Team_Score = 'score2'
    )
```


### Renaming categorical data to make Playoff Rounds more readable.

```{r}
mlb_playoff_games |>
    mutate(
        Playoff_Round = recode(Playoff_Round, 
                         'w' = 'World_Series', 
                         'l' = 'Leage_Championship', 
                         'd' = 'Division_Series', 
                         'c' = 'Wild_Card')
    )
```


## Conclusion: 
#### Objective to use 538 MLB ELO dataset to create a subset of only playoff games was achieved.










